---
title: "Net energy metrics reveal striking disparities across United States household energy burdens"
author: "Eric Scheier"
institute: 
  - "eric@scheier.org"
  - "INFORMS 2021"
  - "This Presentation Was Recorded On:"
date: "`r format(Sys.time(), '%Y-%B-%d')`"
output: 
  beamer_presentation:
    slide_level: 1
    toc: false #true
    theme: "Berkeley"
    colortheme: "seagull"
    fonttheme: "structurebold"
header-includes:
  - \usepackage{booktabs}
  - \usepackage{makecell}
  - \setbeameroption{show notes} #{show/hide notes}
  - \setbeamerfont{note page}{size=\tiny} 
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)

version_path <- "net_energy_equity_major_review_files"

knitr::opts_chunk$set(
  echo=FALSE,
  message=FALSE,
  warning=FALSE
  )
```

# Energy Affordability is Inequitable in the US

```{r preview, fig.cap=""}
knitr::include_graphics(paste(version_path,"figure-latex/continental-map-1.pdf",sep="/"))
```

\note{
Energy inequity in the United States is an issue of increasing urgency. Few policyrelevant datasets evaluate the energy burden of typical American households. Here, we develop a framework using Net Energy Analysis and socioeconomic data at the household level to observe systematic energy inequity across the United States among critical groups that heed policy attention. We find substantial instances of energy poverty in the United States – 16\% of households experience energy poverty. While 94\% of households below the Federal Poverty Line also face energy poverty, more than 5.2 million households above the Federal Poverty Line face this scarcity. We identify that energy expenditures across census tracts disproportionately burden Black, Hispanic, and Native American communities, and we recommend the United States develop a more inclusive federal energy poverty categorization that provides greater assistance for household energy costs.
}

# Energy Burden & Net Energy Return

\[
G = Gross\ Income
\]
\[
S = Spending\ on\ Energy
\]

\[
Energy\ Burden\ (E_{b})= \frac{S}{G}
\]
\[
Household\ Net\ Energy\ Return\ (N_{h}) = \frac{G - S}{S}
\]

\[
E_{B}^{*} = \frac{S}{G} = 6\%
\]
\[
N_{h}^{*} ~16 \Rightarrow Household\ at\ Energy\ Poverty\ Line
\]

\note{
While a variety of thresholds have been developed and explored, energy-poor households in the US are commonly defined in terms of Eb as those with an expenditure of greater than 6\% of household income on energy.

While a helpful place to start, EB has certain drawbacks. A simple proportion does not account for the fact that money spent on energy cannot be spent elsewhere and is therefore not useful income to the household for the purposes of measuring prosperity, effectively depressing the average EB, by definition.

Energy expenditures are a small proportion of even the most impoverished households’ total income, Eb is almost always a very small percentage (<10\%). This leads to issues with interpretability in public discourse and policy settings and may even affect program outcomes that are based on Eb if small numbers are rounded to even the nearest hundredth of a percent. If the household is above an energy poverty line defined by such a metric for public policy purposes, the family may not receive critical support.

Ratios of this type can be useful when delineating across income quantiles or other categories - particularly for vulnerable populations where energy poverty poses a significant difficulty or affordability threshold not captured by measures of absolute poverty. However, Eb is often portrayed at a population scale, which can be skewed by outliers within the population.

Finally, presenting the relationship between household income and energy expenditures as a proportion with income in the denominator suggests that improvements can only be made by decreasing energy expenditures or increasing incomes. However, in reality, there is a positive relationship between energy expenditures and incomes because energy is an input to wealth-creating processes. Households consume energy to unlock the utility that energy services provide to them as participants in society, whether to cook food or connect to the internet. Therefore, a metric describing the efficiency of household wealth creation may be more intuitive with income in the numerator.

Net Energy Return, which describes the newly released potential to do work as a result of some activity, is recommended as a basis for future analysis, especially in the study of macro-energy systems like the US residential housing stock. The main contribution of the study is the application of NER as an indicator of household energy poverty.

Translated into its relative level for Net Energy Return, the 6\% energy poverty line is approximately 16. This means that a household that earns less than approximately \$16 of income for every dollar it spends on secondary energy.
}

# Why Net?

```{r comparison, fig.cap=""}
knitr::include_graphics(paste(version_path,"figure-latex/metric-comparison-1.pdf",sep="/"))
```


\note{
Net Energy Return has several useful mathematical properties: it can smoothly handle systems with negative incomes and energy costs, accept households with zero income, and emphasize extreme incomes and energy costs in an interpretable fashion, as shown in Figure 1. While Eb appears inversely correlated to income, this is primarily driven by a long tail of households with zero or very low incomes, often with energy expenditures exceeding their incomes (approximately n=118,000 households in the dataset have Eb>100\% or Eb<0\%). Due to the structure of the Eb equation (Equation 1), the Eb of these households approaches infinity and cannot be captured on the standard 0-100\% scale the metric is intended to be interpreted within: around 37,000 homes have an infinite energy burden. Since our source data is estimated and provided for the express purpose of exploring low-income communities, we are hesitant to discard these households as outliers. Nh provides a framing of the same dataset that allows for exploration of most households on a similar scale without the long tail. Nh appears positively related to income, and most communities appear within a few orders of magnitude. Utilizing Nh avoids discarding low-income communities as outliers in energy poverty analysis. Furthermore, Nh offers a way to view the relationship between energy expenditures and income such that the wide disparity between those in broader poverty is immediately apparent. Many households with moderate-to-high Ebs are actually higher-income households with high energy expenditures, making their Nhs quite high (e.g., >\$100 of income per \$1 of energy spending). Only households with no energy costs are excluded from the analysis, whereas households with no energy costs or incomes must be excluded from an analysis utilizing Eb.

}

# New Data

Low-Income Energy Affordability Data (LEAD)

**Purpose**: "to help state and local partners understand housing and energy characteristics for the low- and moderate-income (LMI) communities they serve"

  + **Source**: U.S. Department of Energy
  + **Method**: Iterative Proportional Fitting (IPF) of responses from the 5-year American Community Survey (ACS-5)
  + **Geography**: Census tracts for 50 states, D.C., & P.R.
  + **Provides**: Incomes and Energy Costs by Housing Characteristics
  + **Years**: 2016, 2018

\note{
Distinct urban and rural analyses have been performed to describe energy inequity in the US. While limited by geographic and demographic focus and a lack of peer-review, these studies have established the proportion of income (G) spent on energy expenditures (S), or energy burden (Eb), as the standard benchmark for energy poverty in the US today (Equation 1). The US Department of Energy (DOE) significantly improved upon previous methodology by assembling its Low-Income Energy Affordability Dataset (LEAD), which estimates incomes and energy expenditures for most households in the US at a census tract scale.
}

# Summary

```{r load-table, include=FALSE}
print_table <- read_csv("comparison_table.csv")
```


```{r table, results="asis"}
print_table %>%  knitr::kable(booktabs=T, escape = F) %>%
  kable_styling(latex_options="scale_down")
  #kableExtra::kable_styling(latex_options = c("striped","hold_position"))
```

\note{
Table 1 shows a summary of these data for households and their average statistics delineated by their incomes relative to the median income of similar size families in the same metropolitan area or non-metropolitan county, known as Area Median Income (AMI). We can see that high burdens are found mostly in the very-low-income group, with an average Eb of 14\% or Nh of 6.2. Income drives the escape from energy poverty: middle and high-income groups do not spend drastically different amounts on energy (21\% and 63\% more, respectively) but earn 3 and 10 times that of the very low-income group, respectively. An Eb of 14\% means that \$71 per month is spent on energy, which is quite high on a monthly income of \$510. The additional \$15 per month that the next rung of moderate-income groups spends on energy represents a minuscule proportion of their income (1\%).

}

# Geography

```{r pressure, fig.cap=""}
knitr::include_graphics(paste(version_path,"figure-latex/continental-map-1.pdf",sep="/"))
```

\note{
Another contribution of this work is that by using Nh, we can display these data spatially across the US to
explore how different communities are experiencing energy outcomes as in Figure 2 and investigate specific
communities at multiple scales such as census tract, county, state, and regional as in Figure 3. This figure shows a spatial display of energy poverty that includes 5.3 million more households that would not be captured by traditional poverty metrics because their incomes are too low (Eb>100\%) or too high (above the FPL). By processing disparate data sources into a coherent structure and providing a convenient open-source tool for others to do the same, these data can be used in urban planning, public policy, and other relevant contexts. Displayed geospatially in Figure 2, the Black Belt in the American Southeast is visibly perceptible as an area of high burden, indicating that low-Nh follows racial lines. Likewise, border populations and immigrant populated areas in the Southwest have higher burdens, as do Native American lands. High burdens can also be seen in rural Northeast states where heating burdens are high. Nh allows a nuanced view of these widely ranging income dynamics by portraying them on a scale that matches the scope of the issues: areas of high energy burden (close to Nh=0) are visible in orange while highly affluent areas are also visually perceptible as dark blue even though these groups’ average metrics are multiple orders of magnitude apart.
}

# Urban Centers

```{r, fig.cap=""}
knitr::include_graphics(paste(version_path,"figure-latex/inset-map-1.pdf",sep="/"))
```

\note{
Urban inequity results in lower Nh populations not showing up in many dense or gentrified urban areas such
as the San Francisco Bay Area, New York City, and New Orleans, as shown in Figure 3. The pervasiveness
of urban energy poverty in Detroit has been studied extensively and shown to have distinct geographic
boundaries down to the street level19. While some of these conclusions are supported by existing evidence
and literature, they should be confirmed with a rigorous analysis of this dataset using the Nh metric for the
reasons explained in Applying NER to Energy Equity: extremely low-income households are not visible on a
0-100\% scale, and households with no income are often discarded as outliers. Additional analyses should
incorporate additional demographic and household dimensions due to potential disparities within census
tracts since diverse neighborhoods may not be represented accurately by aggregate census tract metrics.
}

# Household Breakdowns

```{r breakdowns, fig.cap=""}
knitr::include_graphics(paste(version_path,"figure-latex/density-charts-1.pdf",sep="/"))
```

\note{
From this high level, we can see in Figure 4.a that approximately 16\% of households in the US experience energy poverty.

Displaying these communities defined by their relationship to the US Federal Poverty Line
(FPL), which indicates income poverty status according to government policy, in Figure 4.b provides a stark picture. While 94\% of households below the FPL also face energy poverty, more than 5.2 million of those households above the FPL face this scarcity, underscoring the relative burden of energy expenditures as a poverty trap and the inadequacy of FPL as an indicator of energy poverty in particular. 

This suggests that energy poverty may be a useful metric for identifying households at risk of other forms of poverty. However, we find that most households experiencing energy poverty are also suffering from a broader lack of access to resources characterized by income-based poverty. Given that the quality of energy used by low-income households is expected to be of similar inherent usefulness, this stark contrast in Nh on the households’ energy investments is surprising.

Figure 4.c shows that households with solar as a primary heating fuel have a higher Nh (Nh=111) than those which rely on other fuel sources (Nh=33), even for households far below the energy poverty line: the average
Nh for energy-impoverished households utilizing solar power is 19, compared to 7 for those relying on any other fuel source. The slope of the Nh density lines in Figure 4.c indicates that these lower-income households seem to experience a different rate of return on Nh from the benefits of increased energy adoption than higher-income households. Why are certain households not receiving the same benefits of their fuel source across the distribution of incomes? This could be due to lower-income households’ low consumption, meaning that the potential savings from implementing energy efficiency measures are lower than for high-income households.
}

# Household Breakdowns

```{r breakdowns-2, fig.cap=""}
knitr::include_graphics(paste(version_path,"figure-latex/density-charts-1.pdf",sep="/"))
```

\note{
Examining these dynamics by the status of homeownership in Figure 4.d reveals further disparities. Though renters and homeowners are similarly distributed below the energy poverty line (28\% of renters and 17\% of homeowners are in energy poverty), there appears to be an advantage of homeownership from a net energy
perspective (Nh=37 for homeowners versus Nh=39 for renters). Only at a relatively high Nh do renters seem to have an advantage: owners of multi-family apartments earn 2.4 times as much as renters of single-family homes when normalized by energy expenditures. Renters face systemic disadvantages in the energy transition;
they typically pay the home’s energy costs while the landlord controls infrastructure upgrades, commonly understood as the split incentive problem36. Tenure matters for more than just equity itself: renters are less likely to take actions to improve their Nhs due to a lack of property rights and split incentives. Even
when action is taken to improve the energy efficiency of a rental building, tenants are less likely to see any economic benefits from it.

Furthermore, Figure 4.e shows that Nh varies widely by racial demography. Asian households have the highest Nh across the entire population distribution (Nh=65), and Indian households have the lowest (Nh=18), with Black households a close second from the lowest (Nh=26). These relative positions are the same across the
entire population distribution, with only White (Nh=38) and Hispanic (Nh=36) populations showing different relative Nhs across the population. Households in communities of color experience energy poverty at a rate 60\% more than those in white communities.

Education level also seems to be correlated with disparate Nh outcomes according to Figure 4.f, with a wide gap between those households in areas with mostly high-school (Nh=25) or college-educated (Nh=40) populations.
}

# State Comparison

```{r states, fig.cap=""}
knitr::include_graphics(paste(version_path,"figure-latex/state-violin-1.pdf",sep="/"))
```

\note{
Assessing the Nhs among different states in Figure 5 presents a counterintuitive picture of how states address energy poverty and energy equity. Nh can equate communities that experience high energy costs and low incomes with those with high incomes and even higher energy costs. This explains why states such as
Connecticut (Nh=26) and Vermont (Nh=23), where 47\% and 30\% higher than average electricity prices may pose affordability threats for communities affected by higher prices, are similarly positioned on the list to states such as Mississippi (Nh=22) and Alabama (Nh=24), which have significant low-income populations
and low per-unit energy prices (22\% and 8\% lower than average, respectively"‘). Not only are households in these states falling behind in terms of income, but net incomes are lower relative to energy expenditures than neighboring states and other parts of the country. This may be appropriate: while the equity issues in
Southeastern states are well studied, states such as Maine that continue to utilize residential heating energy sources like oil and fuelwood may suffer not only from lack of efficiency but also health impacts. States may need to pay attention to these dynamics from an affordability perspective, and further targeted energy assistance may be needed based on new metrics.

Although they are the states with the highest Nh, California (Nh=59) and Colorado (Nh=63) are not immune to these problems and likely represent a greater spread and diversity of energy affordability impacts. Likely,
this diversity captures the benefits accrued by early adopters and the challenges of having high populations of those struggling with energy poverty and high housing costs. In many of these places, residents have self-sorted into geographic areas based on the overall costs of living. Also, advances in clean energy legislation are a common thread among the top-performing states on an Nh basis, signaling the value of strong decarbonization targets and accompanying policies to ensure electricity affordability for low-income households.

Visualizing the proportion of end-use energy sourced from the combustion of fossil fuels in Figure 5 shows that such reliance does not necessarily lead to a more affordable system for energy consumers. Households in states with a high proportion of fossil fuel are no less likely to have high Nhs than those in other states that rely on clean energy, defying the conventional wisdom that fossil fuel consumption is a chosen tradeoff between environmental health and affordability for citizens.
}

# Conclusions

+ 16% of households experience energy poverty as presently defined (>6% of household income on energy expenditures)
+ 94% of households below the federal poverty line also face energy poverty
+ More than 5.2m of those above the federal poverty line face this scarcity
+ Fossil fuel reliance does not lead to lower energy poverty rates at a state scale
+ Solar power is associated with decreased energy poverty at a household scale
+ Energy poverty presents an obstacle to decarbonization and wealth creation

# Conclusions

Energy expenditures in the US disproportionately burden those in:

+ Communities of color
+ Communities with limited education
+ Rural cold areas
+ Urban Centers

The United States should develop and implement a federal energy poverty line.

# Thank You!

```{r review, fig.cap=""}
knitr::include_graphics(paste(version_path,"figure-latex/continental-map-1.pdf",sep="/"))
```

eric@scheier.org

https://github.com/ericscheier/net_energy_equity
